---
title: "Project"
output: pdf_document
---

Simulating the exponential

```{r}
n= 40
lambda = 0.2
k=10000
list_of_exponential = array(1:k)

for(i in 1:k){
  list_of_exponential[i] = mean(rexp(n, lambda))
}

```

Question 1- Let's calculate where the distribution of means is centered and compare to the theoretical center

```{r}
mean_of_average = mean(list_of_exponential)
mean_of_average
```
Meanwhile, the expected value for mean is 1/lambda = 5. 
```{r}
sd_mean = sd(list_of_exponential)
sd_mean

```
The expected Standard Deviation  is {[1/(lambda)^(2)]/n}^(0.5)
```{r, echo=FALSE}
std<-((1/lambda)/sqrt(40))
std
```


1- When we compare the theoretical center we were expecting the average to be 5. it's 5.07.

2- And we were expecting the standard deviation to 0.79 and it's 0.784.


3-For the approximation to a normal, we can check the figure 1 in the appendix. the curve is a normal(5,0.7906) and the histogram is referent to the the mean distribution.
We can see that it clearly looks like a normal. To understand how well fitted is the model I have generate 1000 normal variables (Figure 2) as if I wanted to check how they look like, and they are as good as the exponential distribution (even better in my opinion). So we can conclude that when n increases, the exponential converges to a normal.

4- As 1/lambda is the mean for this distribution we have that at 95% of confidence the mean is in this interval:
```{r, echo=FALSE}
mean(list_of_exponential) + c(-1,1)*1.96*sd(list_of_exponential)/sqrt(nrow(list_of_exponential))

```

#Conclusion:

From the points above we can conclude that the hypothesis of convergence is consistent for exponential distribution and the mean of n exponential converges to the distribution mean.


\newpage

#Appendix

Figure 1

```{r, fig.width=8, fig.height=4.0, echo=FALSE}
    hist(list_of_exponential, freq= FALSE, col = "blue", main = "Histogram of simulated mean", xlab = "mean of exponential")
    curve(dnorm(x, mean=5, sd=std), 
          col="red", lwd=2, add=TRUE)


```


Figure 2

```{r, fig.width=8, fig.height=4.0, echo=FALSE}
x<-rnorm(1000,5, std) 
hist(x, freq= FALSE, col = "green",main = "Histogram of normal with mean(5,0.7906)")
curve(dnorm(x, mean=5, sd=std), 
          col="red", lwd=2, add=TRUE)

```