---
title: "Week 10 Application Part: Graph Analytics with R"
output: beamer_presentation
---
```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, results='hide'}
library("knitr")
options(width = 1000)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 200)
```


## Preparation
```{r, warning=FALSE, results='hide'}
library("igraph") # graph analytics library in R
```

## Creating Graphs 

*From an edge list*

```{r, warning=FALSE}
g <- graph( c(1,2, 1,5, 2,3, 2,5, 2,4,
              3,4, 4,5), n=5, 
            directed=F)
```

## Creating Graphs

*Create complete or null graphs*

```{r, warning=FALSE}
full_g <- graph.full(5)
empty_g <- graph.empty()
```

## Show Nodes and Edges

```{r, warning=FALSE}
V(g)
V(full_g)
V(empty_g)
```

## Show Nodes and Edges

```{r, warning=FALSE}
E(g)
```

## Read Graph Files

```{r, warning=FALSE}
karate <- read.graph("http://cneurocvs.rmki.kfki.hu/igraph/karate.net", 
                     format="pajek")
```

## Show Degree Distribution

```{r fig.width = 4, fig.height= 3}
hist(degree(karate))
```
## Check Betweenness Values

```{r, warning=FALSE}
betweenness(g)
edge.betweenness(g)
```

## Visualization of Graphs

```{r fig.width = 4, fig.height= 3}
plot(g, vertex.label=V(g), vertex.size=0)
```

## Visualization of Graphs

```{r fig.width = 6, fig.height= 4}
plot(karate, vertex.label=V(karate), vertex.size=0)
```

## Cluster a Graph

*Check the connected components*

```{r, warning=FALSE}
is.connected(karate)
clusters(g)
```

## Cluster a Graph

*Cluster karate club network*

```{r, fig.width = 6, fig.height= 4}
wc <- edge.betweenness.community(karate)
plot(wc, karate)
```

## Lab Preparation

sample_graph <- barabasi.game(100, directed=F)

## Lab Problems:

- Create a trivial graph with 3 nodes A,B and C. There is an edge between A and B. There is an edge between B and C. Plot this graph in the end.
- Show the degree distribution and plot the histogram of the degrees.
- Show the correlation between betweenness and degree for each node.
- Partition the sample graph and show its components.


