---
title: "Week 3 Application Part"
output: beamer_presentation
---


## Identification of Missing Values
```{r echo=TRUE}
a <- cbind(x=c(1,2,NA), y=c(NA,1,2))
a
complete.cases(a)
```

## Identification of Missing Values
```{r echo=TRUE}
a <- cbind(x=c(1,2,NA), y=c(NA,1,2))
a[!complete.cases(a),]
```


## Removing Instances with Missing Values
```{r echo=TRUE}
a <- cbind(x=c(1,2,NA), y=c(NA,1,2))
na.omit(a)
```


## Changing Missing Values
```{r echo=TRUE}
a <- cbind(x=c(1,2,NA), y=c(NA,1,2))
a[is.na(a)] <- 0
a
```


## Creation of Test and Training Data

Create random split
```{r echo=TRUE}
rn_train <- sample(nrow(iris), floor(nrow(iris)*0.7))
train <- iris[rn_train,]
test <- iris[-rn_train,]
```

## Creation of Test and Training Data

Cross validation:

```{r echo=TRUE, }
library(caret)
library(mlbench)
data(Sonar)
folds <- createFolds(Sonar$Class)
```


## Creation of Test and Training Data

Cross validation:

```{r echo=TRUE, }
str(folds)
```


## Creation of Test and Training Data

Cross validation:
```{r echo=TRUE, }
for (f in folds){
  train <- Sonar[-f,] 
  test <- Sonar[f,]
  #do stuff
  }
```

## Lab Section - 1

1. Find the instances in airquality dataset with missing values.
2. Remove instances with missing data in airquality dataset.
3. Randomly split mtcars dataset to 80% training samples and 20% test samples.
4. Create 10 fold cross validation pairs for iris dataset.

## Lab Section - 2

Find the following probabilities in mtcars dataset (hint: use prop.tables from week 2)

1. P(4cylinder | 3 gears)
2. P(4cylinder,4 gears)
3. P(4cylinder)

## Homework 

1. Individual work
2. Send code+report through blackboard.
3. No paper report required
4. Deadline: 22 Feb 2015 (hard)

