---
title: "Week 4 Application Part"
output: 
  beamer_presentation:
    fig_width: 4
    fig_height: 4
---



## Finding Correlation Coefficients
Spearman:
```{r echo=TRUE}
cor(iris$Sepal.Length, iris$Petal.Length, 
    method="spearman")
```
Pearson:
```{r echo=TRUE}
cor(iris$Sepal.Length, iris$Petal.Length, 
    method="pearson")
```

## Correlation Matrix
```{r echo=TRUE}
cor(iris[,1:3])
```

## Visualizing Correlation
```{r echo=FALSE, eval=TRUE}
require("corrplot")
```

```{r setup, echo=FALSE}rel
library("knitr")
opts_chunk$set(dev = 'pdf')
```

```{r echo=TRUE, fig.width=10, fig.height=10}
corrplot(cor(mtcars), method="ellipse")
```

## Significance of Correlation Coefficients
```{r echo=TRUE}
cor.test(iris$Sepal.Length, iris$Petal.Length, 
         method = c("pearson"))
```

## Linear Regression: Fitting the Model
```{r echo=TRUE}
plot(airquality$Wind, airquality$Temp)
```

## Linear Regression: Fitting the Model
```{r echo=TRUE}
model_ulm <- lm(Wind~Temp, data=airquality) 
summary(model_ulm)
```
## Linear Regression: Prediction
```{r echo=TRUE, eval=FALSE}
rn_train <- sample(nrow(airquality), 
                   floor(nrow(airquality)*0.8))
train <- airquality[rn_train,c("Wind","Temp")]
test <- airquality[-rn_train,]
model_ulm <- lm(Wind~Temp, data=train) 
prediction <- predict(model_ulm, interval="prediction", 
                      newdata =test)
```
```{r echo=FALSE, eval=TRUE}
rn_train <- sample(nrow(airquality), 
                   floor(nrow(airquality)*0.8))
train <- airquality[rn_train,c("Wind","Temp")]
test <- airquality[-rn_train,]
model_ulm <- lm(Wind~Temp, data=train) 
prediction <- predict(model_ulm, interval="prediction", 
                      newdata =test)
```

## Linear Regression: Error Distribution
```{r echo=TRUE, fig.width=10, fig.height=10}
errors <- prediction[,"fit"] - test$Wind
hist(errors)
```

## Linear Regression: RMSE
```{r echo=TRUE}
sqrt(sum((prediction[,"fit"] - test$Wind)^2)/nrow(test))
```

## Linear Regression: PRED(25)
Find the percentage of cases with less than 25 percent error:
```{r echo=TRUE}
rel_change <- 1 - ((test$Wind - abs(errors)) / test$Wind)
table(rel_change<0.25)["TRUE"] / nrow(test)
```

## Lab
**Preparation**
Required Libraries
```{r echo=TRUE, eval=FALSE}
install.packages("corrplot")
require("corrplot")
```
## Lab
**Preparation**
Data load
```{r echo=TRUE, eval=FALSE}
library(RCurl)
u <- getURL("http://vincentarelbundock.github.io/Rdatasets/csv/Ecdat/Computers.csv")
c_prices <- read.csv(text = u)
```

## Lab Questions
1- Find spearman correlation between hard disk space and ram.

2- Visualize the correlation of the numeric columns in the computer prices dataset.

3- Choose a single variable to predict price and build an univariate linear regression model.

4- Experiment with 30 percent split of the data. Report error distribution, RMSE and pred(25)