---
title: "Week 7 Application Part"
output: beamer_presentation
---
```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, results='hide'}
library("knitr")
options(width = 1000)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 300)
```


## Preparation
```{r, warning=FALSE, results='hide'}
library("RWeka") # for datasets
library("ROCR") # visualize performance of classifiers
library("caret") # for confusion matrix
library("e1071") # may be needed for caret
diabetes <- read.arff(system.file("arff", "diabetes.arff", 
                                  package = "RWeka"))
```

## Solution of the Optional Problem From Last Week

In this part, we will go through the R code steps based on the 
solution for lab 6.

## Dynamically Create R Formulas

You do not need to write variable names every time on regression models.

```{r}
formula_text <- paste(names(diabetes[9]),"~",
                paste(names(diabetes[1:8]), collapse="+")
                )
formula <- as.formula(formula_text)
formula
```

## Fit a logistic Regression Model

*Remember: fit checks your model on the same data.*

```{r}
diabetes$class <- as.integer(diabetes$class) - 1
fit <- glm(formula,data=diabetes,family=binomial())
summary(fit) # display fit results
```

## Predict using Random Split

```{r}
rn_train <- sample(nrow(diabetes), 
                   floor(nrow(diabetes)*0.7))
train <- diabetes[rn_train,]
test <- diabetes[-rn_train,]
fit <- lm(formula, data=train) 
```

## Predict using Random Split

```{r}
test$scores <- predict(fit, type="response", 
                      newdata=test)

pred<-prediction(test$scores, test$class)
```

## Create a Confusion Matrix

```{r}
c <- confusionMatrix(as.integer(test$scores > 0.5), 
                     test$class)
c$table
```
## Plot the ROC Curve
```{r, fig.width = 4, fig.height= 3}
perf<-performance(pred,"tpr","fpr")
plot(perf, lty=1)
```

## Using rWeka For Logistic Regression
```{r}
diabetes <- read.arff(system.file("arff", "diabetes.arff", 
                                  package = "RWeka"))
weka_fit <- Logistic(formula, 
                    data = diabetes)
evaluate_Weka_classifier(weka_fit, numFolds = 10)
```
## Lab Preparation

```{r, warning=FALSE, eval=FALSE}
library("RWeka") # rweka (embedded Weka software)

diabetes <- read.arff(system.file("arff", "diabetes.arff", 
                                  package = "RWeka"))

```

## Lab Problems:
*Using standard mtcars dataset*

1. Create a formula am~hp+mpg+gear+carb. 

2. Create a logistic regression model using the parameters.

3. Check the performance of the logistic regression model and show ROC curve and confusion matrix.

4. Using rweka interface compare the performance of knn with logistic regression for the same problem.

5. (Optional) Play with the model parameters and attributes to increase the performance of your model.

