---
title: "Week 8 Application Part"
output: beamer_presentation
---
```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F, results='hide'}
library("knitr")
options(width = 1000)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 300)
```


## Preparation
```{r, warning=FALSE, results='hide'}
library("tm") # text mining package for R
library("RCurl") # get text from the web
library("klaR") # implementation of naive bayes in R
library("caret") # experimental design
```

## Importing and Analyze Text Data
```{r, warning=FALSE}
data("crude")
# strip white space
crude <- tm_map(crude, stripWhitespace)
# remove stop words
crude <- tm_map(crude, removeWords, stopwords("english"))
# stem words
crude <- tm_map(crude, stemDocument)
```

## Finding Frequent terms

```{r, warning=FALSE}
dtm <- DocumentTermMatrix(crude)
# at least 10 occurrence
findFreqTerms(dtm,10)
```
## Finding associations

Finding associations

```{r, warning=FALSE}
findAssocs(dtm, "opec", 0.8)
```

## Finding tf

```{r, warning=FALSE}
dtm <- DocumentTermMatrix(crude, 
       control = list(weighting = weightTfIdf))
# three important terms for the first document
sort(as.matrix(dtm)[1,], decreasing=T)[1:3]
```

## Naive Bayes Implementation 

```{r}
rn_train <- sample(nrow(iris), 
                   floor(nrow(iris)*0.7))
train <- iris[rn_train,]
test <- iris[-rn_train,]
model <- NaiveBayes(Species~., data=train)
```

## Naive Bayes Implementation 
Show posterior probability densities
```{r fig.width = 4, fig.height= 3}
plot(model)
```
## Naive Bayes Implementation 

```{r, warning=FALSE, fig.width = 4, fig.height= 3}
# make predictions
predictions <- predict(model, test)
# summarize results
confusionMatrix(predictions$class, predictions$class)
```

## Lab Preparation

```{r, warning=FALSE, eval=FALSE}
library("RWeka") # rweka (embedded Weka software)

diabetes <- read.arff(system.file("arff", "diabetes.arff", 
                                  package = "RWeka"))
data("acq")

```

## Lab Problems:

1. Process acq text by stemming and create a document term matrix.
2. Show 3 important terms for **each** document using tf . 
2. Run Naive Bayes using diabetes dataset.
3. Show posterior probability densities of diabetes.
4. (optional) Find and use the naive bayes functionality of weka for the same dataset.

