---
title: "Handwriting Recognition Using KNN"
author: "Andresa de Andrade"
date: "March 26, 2015"
output: pdf_document
---
\newpage

#Introduction
One of most challenging problems in these days is to recognize patterns for images.
In this document the methodology KNN is applied in order to find patterns in the hand writing of numbers from 0-9 and try to predict them using these patterns.

In this aspiration we have challenges such as: 
- each person has their own way of writing a number;
- some numbers are very similar to others, like 1 and 7, or 6, 8 and 9.
- find the better number of neighborhood in order to optimize the algorithm.

#Data Explanation
In order to apply this algorithm we have each number divided in a matrix 4 x 4 resulting in 16 variables presented as columns in our data set. The last column is the number class.

For this project we have two files, one for training and one for testing. The testing data set contains 3497 samples and the training has 7493.

#KNN

So training our model to use k=5, our confusion matrix has too many 7 classified as number 1and too many 6 as number 0.

```{r, echo=FALSE}
library(RWeka)
tdata<-read.csv("pendigits.tra.csv", h=F)
testing_data<-read.csv("pendigits.tes.csv", h=F)
vector_names<-c("x1")
for(i in 2:16){
  vector_names<-rbind(vector_names, paste("x",i, sep=""))

}
vector_names<-rbind(vector_names, "class")
vector_names<-as.vector(vector_names)
names(tdata)<-vector_names
names(testing_data)<-vector_names
testing_data$class<-as.factor(testing_data$class)
tdata$class<-as.factor(tdata$class)
classifier <- IBk(class ~., data = tdata,
                  control = Weka_control(K = 5))
evaluate_Weka_classifier(classifier, testing_data)


```
So we have an accuracy of 97.59% of the cases, and in this project we need to find the best K for set of data.

Running the same algorithm for other cases we found that k=1 has the best Correctly Classified Instances:

```{r, echo=FALSE}

classifier <- IBk(class ~., data = tdata,
                  control = Weka_control(K = 1))
evaluate_Weka_classifier(classifier, testing_data)


```

So running the algorithm for k=1 we have a percentage of accuracy of 97.74% of the cases. We also reduce the cases that 7 and 1 have gotten mislabeled.